---
title: "ESWD Lab1"
author: "JLewyckyj"
date: "10/25/2018"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```


#Problem 1

```{r}
setwd("/nfs/home/jsl2985/z/Everything Starts with Data/Lab1")
library(expm)
```

##1a

```{r}
markov <- read.table("markov100.txt")
M <- as.matrix(markov)
a <- c(1, integer(99)) #Create a vector of 1 and consecutive zeroes (probability distribution of initial state at time 0)
ans1 <- a %*% (M%^%10) #Multiply the vector by the matrix 10 times (10 transitions)
print(ans1[5]) #Print the probability of being in State 5 after 10 transitions
```

Here, we read in the table and convert it to a matrix. We create the probability distribution of initial state at time 0 (1,0,0,...), and multiply that vector by the matrix 10 times for 10 transitions, and print the probability of being in state 5.

##1b

```{r}
b <- c(rep(1/3,3), integer(97)) #Create a vector of 1/3, 1/3, 1/3 and consecutive zeroes (probability distribution of initial state at time 0)
ans2 <- b %*% (M%^%10) #Multiply the vector by the matrix 10 times (10 transitions)
print(ans2[10]) #Print the probability of being in State 10 after 10 transitions
```

Here, we create the probability distribution of initial state at time 0 (1/3, 1/3, 1/3, 0, 0, 0, ...), and multiply that vector by our matrix 10 times for 10 transitions, and print the probability of being in state 10.

##1c

```{r}
Q <- t(M) - diag(100) #Transpose the matrix, and subtract identity matrix of size 100
Q[100,] <- c(rep(1,100)) #Replace last row with all 1s
rhs <- c(rep(0,99),1) #Create vector of 99 zeroes and 1 one
Pi <- solve(Q) %*% rhs #Multiply this vector by the inverse of Q
print(Pi[1]) #Print steady-state probability of being in State 1
```

Here, we transpose the initial matrix, and subtract an identity matrix of size 100, and replace the last row with all 1s. We then create a vector of 99 0s and one 1, and multiply that vector by our new matrix, and print the steady-state probability of being in state 1.

##1d

```{r}
B <- M[1:99, 1:99] #Get rid of last row from the matrix
Q2 <- diag(99) - B #Subtract matrix from identity matrix
e <- c(rep(1,99)) #Create vector of 99 ones
m <- solve(Q2) %*% e #Multiply vector by inverse of matrix Q2
print(m[1]) #Print mean passage time from State 1 to State 100
```

Here, we remove the last row of the matrix, since that is where we are transitioning to. We subtract this matrix by an identity matrix of size 99. Then, we create a vector of 99 ones, and multiply that vector by the inverse of our new matrix, and print the mean passage time from state 1.

#Problem 2

##A

```{r}
webtraffic <- read.table("webtraffic.txt", header=TRUE)

w2 <- as.matrix(webtraffic)

traffic <- matrix(colSums(w2), 9, 9, byrow=TRUE) #Use colSums to create a vector with the total traffic between each pair of states. Re-align this vector into a new 9x9 matrix

print(traffic)
```

We use colSums to create a vector with the total traffic between each pair of states. Then, we re-align this vector into a new 9x9 matrix.

##B

```{r}
traffic[9,1] <- 1000 

P <- traffic / rowSums(traffic) #Divide each cell in the matrix by the sum of each row to get the one step transition probability matrix

print(P)
```

Here, we divide each cell in the matrix by the sum of each row to turn the raw traffic numbers to the 1-step transition probability matrix.

##C

```{r}
Q <- t(P) - diag(9) #Transpose the matrix, and subtract identity matrix of size 9
Q[9,] <- c(1,1,1,1,1,1,1,1,1) #Replace last row with all 1s
rhs <- c(0,0,0,0,0,0,0,0,1) #Create vector of 8 zeroes and 1 one
Pi <- solve(Q) %*% rhs #Multiply this vector by the inverse of Q

print(Pi)
```

Here, we transpose the 1-step transition probability matrix, and subtract an identity matrix of size 9, and replace the last row with all 1s. We then create a vector of 8 zeroes and 1 one, and multiply this vector by the inverse of the matrix we calculated previously.

##D

```{r}
avg_time <- c(0.1, 2, 3, 5, 5, 3, 3, 2)

B <- P[1:8,1:8]
Q <- diag(8) - B
m <- solve(Q) %*% avg_time
print(m[1])
```

Here, we remove the last row of the matrix, since that is where we are transitioning to. We subtract this matrix by an identity matrix of size 8. Then, we create a vector for average time spent on each page, and multiply that vector by the inverse of our new matrix, and print the mean passage time from state 1.

##E

```{r}
traffic[2,3] <- 230 * 0.7
traffic[2,4] <- 321 * 0.8
traffic[2,6] <- 230 * 0.3
traffic[2,7] <- 321 * 0.2

P2 <- traffic / rowSums(traffic) #Divide each cell in the matrix by the sum of each row to get the new one step transition probability matrix

Q <- t(P2) - diag(9) #Transpose the matrix, and subtract identity matrix of size 9
Q[9,] <- c(1,1,1,1,1,1,1,1,1) #Replace last row with all 1s
rhs <- c(0,0,0,0,0,0,0,0,1) #Create vector of 8 zeroes and 1 one
Pi2 <- solve(Q) %*% rhs #Multiply this vector by the inverse of Q

print(Pi2)

var(Pi) #Compare variances
var(Pi2) #Compare variances
```

Here, we re-assign values in the traffic matrix, and then repeat the steps from Part C. It looks the variance slightly decreased in the Pi vector, though states 3 and 4 are still overrepresented compared to states c(2, 6:8).
