---
title: "ESWD Lab2"
author: "JLewyckyj"
date: "11/10/2018"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r}
setwd("/nfs/home/jsl2985/z/Everything Starts with Data/Lab2")
library(caret)

redwine <- read.table("redwine.txt", header=TRUE)
```

#1

```{r}
mean(redwine$RS, na.rm=TRUE)
mean(redwine$SD, na.rm=TRUE)
```

The mean of RS (Residual Sugar) is 2.538.
<br>
The mean of SD (Total Sulfur Dioxide) is 46.298.

#2

```{r}
#Create dataset without NAs for SD
redwine.obs <- subset(redwine, !is.na(redwine$SD))
```

The SD and FS columns for redwine.obs are the vectors for the data without missing values for SD.

```{r}
ABC <- lm(SD ~ FS, redwine.obs)
summary(ABC)
coefficients(ABC)
```

The coefficients for the model are 13.19 for B0 and 2.09 for B1.

#3

```{r}
#Create dataset with only rows for missing values of SD
redwine.missing <- subset(redwine, is.na(redwine$SD))

#Predict the missing values of SD using values of FS
redwine.missing$SD.predictions <- predict(ABC, redwine.missing)

print(redwine.missing$SD.predictions)
```

```{r}
#Impute SD.predictions for the missing values of SD

redwine.missing$SD <- redwine.missing$SD.predictions #Assign predictions column to SD
redwine.missing <- redwine.missing[,1:12] #Remove predictions column

#Re-combine datasets

redwine.new <- rbind(redwine.obs, redwine.missing)

#New average of SD

mean(redwine.new$SD)
```

The new mean for SD is 46.302.

#4

```{r}
#Average value imputation function

avg.imp <- function (a, avg){
        missing <- is.na(a)
        imputed <- a
        imputed[missing] <- avg
        return (imputed)
}

RS.avg <- mean(redwine.new$RS, na.rm=TRUE)

redwine.new$RS <- avg.imp(redwine.new$RS, RS.avg) #Impute missing values with the average for RS

mean(redwine.new$RS)
```

The mean remains the same for RS at 2.538.

#5

```{r}
winemodel <- lm(QA ~ ., redwine.new) #Multiple regression model

winemodel$coefficients #Print out coefficients
```

B0: 47.20
<br>
B1 for FA: 0.068
<br>
B2 for VA: -1.097
<br>
B3 for CA: -0.179
<br>
B4 for RS: 0.0259
<br>
B5 for CH: -1.631
<br>
B6 for FS: 0.0035
<br>
B7 for SD: -0.0029
<br>
B8 for DE: -44.817
<br>
B9 for PH: 0.036
<br>
B10 for SU: 0.945
<br>
B11 for AL: 0.247

#6

```{r}
summary(winemodel)
```

PH is least likely to be related to QA on account of it's 0.414 p-value, not significant at any level.

#7

```{r}
# Define training control
set.seed(123) 
train.control <- trainControl(method = "cv", number = 5)
# Train the model
cv_model <- train(QA ~ ., data = redwine.new, method = "lm",
               trControl = train.control)
# Summarize the results
print(cv_model)

summary(cv_model)
```

```{r}
#Use the cross-validated model to predict QA in the redwine.new dataframe

redwine.new$predictions <- predict(cv_model, redwine.new) #Create predictions column
redwine.new$error <- abs(redwine.new$QA - redwine.new$predictions) #Define error as the absolute value of the difference between QA and the prediction

mean(redwine.new$error)
```

The average absolute value of the error between QA and the prediction given by the cross-validated model is 0.501.

#8

```{r}
mPH <- mean(redwine.new$PH)
print(mPH)

sdPH <- sd(redwine.new$PH)
print(sdPH)
```

The mean of PH is 3.306.
<br>
The standard deviation of PH is 0.392.

```{r}
t1 <- mPH - 3*sdPH #Set thresholds for what should be removed
t2 <- mPH + 3*sdPH

redwine2 <- subset(redwine.new, PH > t1 & PH < t2) #Keep only observations within the threshold for PH

dim(redwine2)
```

There are 1580 rows, so 19 were removed due to outliers in PH.

#9

```{r}
redwine2 <- redwine2[,c(1:12)] #Remove prediction and error columns from the CV model

winemodel2 <- lm(QA ~ ., redwine2)
summary(winemodel2)
```

This model is likely better than the previous one, with a higher F-statistic (81.21 vs. 80.6), and a higher R-squared (0.3629 vs. 0.3584).

Based on p-values, the 5 most likely attributes to be related with QA are: <br>
1: AL / Alcohol (p-value < 2e-16) <br> 
2: VA / Volatile Acidity (p-value < 2e-16) <br>
3: SU / Sulphates (p-value = 3.46e-15) <br>
4: CH / Chlorides (p-value = 6.60e-06) <br>
5: SD / Total Sulfur Dioxide (p-value = 2.16e-05) <br>
